\documentclass[../main.tex]{subfiles}

\begin{document}
\chapter{Additive and abelian categories}
We will set sheaves aside for the moment and layout the algebraic foundations for the lectures to come, in which we will study the field of \emph{homological algebra}.
In this lecture, we define \emph{abelian categories}.
The idea is to axiomatise the defining properties of the category of modules over a ring.
As a result, categories of \(R\)-modules for some ring \(R\) will be abelian.
Strikingly, this axiomatisation is `strict' in the sense of the Freyd--Mitchell theorem (which will not be treated in this course), which says that any small abelian category can be embedded in some category of modules~\cite[Theorem~1.6.1]{WeibelHomologicalAlgebra}. 

\section{Additive categories}
\remyquote{We can only write down diagrams, prove they commute, and run away as fast as we can.}
\begin{defn}
A \emph{pre-additive} category $\cat C$ is a locally small category where each hom-set $\Hom_{\cat C}(X, Y)$ is endowed with the structure of an abelian group such that composition
\[ \Hom[\cat C](Y,Z)\times\Hom[\cat C](X,Y) \to \Hom[\cat C](X,Z)\text{,} \quad (g,f)\mapsto gf \]
is bilinear.
\end{defn}

Expanding the definitions, this means in particular that between every two objects \(X\) and \(Y\) in a pre-additive category, there is a zero map \(0\colon X\to Y\), and that two parallel maps \(f,g\colon X\to Y\) have a sum \(f+g\colon X\to Y\).

The following lemma shows that in a pre-additive category, finite coproducts and finite products coincide.

\begin{lem}\label{lem:pre-additive-categories-are-nice}
Let $\cat C$ be a pre-additive category.
\begin{enumerate}
  \item If $X \in \cat C$, then the following are equivalent:
  \begin{enumerate}
    \item\label{lem:pre-additive-categories-are-nice:initial} $X$ is initial;
    \item\label{lem:pre-additive-categories-are-nice:terminal} $X$ is terminal;
    \item\label{lem:pre-additive-categories-are-nice:identity-zero} $\id_X = 0 \in \Hom_{\cat C}(X,X)$;
  \end{enumerate}
  \item If $X, Y, Z \in \cat C$ then the following are equivalent:
  \begin{enumerate}
    \item\label{lem:pre-additive-categories-are-nice:coproduct} There are maps \[X \xrightarrow{i} Z \xleftarrow{j} Y\] making $Z$ the coproduct.
    \item\label{lem:pre-additive-categories-are-nice:product} There are maps \[Z \xrightarrow{p} X \xleftarrow{q} Y\] making $Z$ the product.
    \item\label{lem:pre-additive-categories-are-nice:biproduct} There are maps \[X \xrightleftarrows{i}[p] Z \xleftrightarrows{j}[q] Y\] such that $p \circ i = \id_X$ and $q \circ j = \id_Y$ and $i\circ p + j \circ q = \id_Z$.
  \end{enumerate}
\end{enumerate}
\end{lem}

\begin{defn}
A \emph{biproduct} of two objects \(X\) and \(Y\) in a pre-additive category is an object $Z$ with maps \[X \xrightleftarrows{i}[p] Z \xleftrightarrows{j}[q] Y\] satisfying $pi = \id_X$, $qj = \id_Y$ and $ip + jq = \id_Z$.
\end{defn}
\begin{rmk}
  Note that these force $q \circ i$ and $p \circ j$ to be zero. To see this it suffices to show that $j \circ q \circ i = 0$, because $j$ is monic.
  This holds because $jqi = (\id*_Z - ip)i = i - ipi = i - i = 0$.
\end{rmk}

\begin{proof}[name={of \cref{lem:pre-additive-categories-are-nice}}]
  \begin{enumerate}
      \item If $X$ is initial, then $\Hom_{\cat C}(X, X) = 0$ so $\id_X = 0$.
    Conversely, if $\id_x = 0$ then every map $f\colon X \to Y$ i n $\cat C$ satisfies $f = f \circ \id_X = f \circ 0 = 0$. So $X$ is initial. This proves \cref{lem:pre-additive-categories-are-nice:initial} is equivalent to \cref{lem:pre-additive-categories-are-nice:terminal}, and it follows dually that \cref{lem:pre-additive-categories-are-nice:terminal} is equivalent to \cref{lem:pre-additive-categories-are-nice:identity-zero}.
    \item If \[X \xrightleftarrows{i}[p] Z \xleftrightarrows{j}[q] Y\] is a biproduct, we saw $qi = 0$ and $pj = 0$. Then $X \xrightarrow{i} Z$ and $Y \xrightarrow{j} Z$ is a coproduct. If \[X \xrightarrow{f} W \xleftarrow{g} Y\] is any cocone, set $h = fp + gq\colon Z \to W$. Then $hi = (fp + gq)i = fpi + gqi = f \circ \id_X + g \circ 0 = f$, and likewise $hj = g$. If $h'\colon Z \to W$ satisfies $h'i = f$ and $h'j = g$ then $h' = h' \circ \id*_Z = h' \circ (ip + jq) = h'ip + h'jq = fp + gq = h$. So \[X \xrightarrow{i} Z \xleftarrow{j} Y\] form a coproduct. Conversely, if these maps form a coproduct, then the cocones 
    \begin{align*}
          X \xrightarrow{\id} X \xleftarrow{0} Y\text{,} && X \xrightarrow{0} Y \xleftarrow{\id} Y      
    \end{align*}
    define maps $p\colon Z \to X$, $q\colon Z \to Y$ with $pi = \id_X, pj = 0, qi = 0, qj = \id_Y$. Then $ip + jq\colon Z \to Z$ satisfies $(ip + jq)i = ipi + jqi = i \circ \id_X + j \circ 0 = i$, and likewise $(ip + jq)\circ j = \ldots = j$. So $ip + jq = \id_Z$ by the universal property of the coproduct. This proves that \cref{lem:pre-additive-categories-are-nice:coproduct} is equivalent to \cref{lem:pre-additive-categories-are-nice:biproduct}, and the equivalence of \cref{lem:pre-additive-categories-are-nice:product} and \cref{lem:pre-additive-categories-are-nice:biproduct} follows dually. \qedhere
    \end{enumerate}
\end{proof}
\begin{defn}
  An \emph{additive} category is a pre-additive category with finite products.
\end{defn}
\begin{rmk}
  By \cref{lem:pre-additive-categories-are-nice}, if $\cat C$ is additive then:
  \begin{itemize}
      \item the terminal object is also initial, hence a zero object $0$;
    \item if $X, Y \in \cat C$ then the biproduct \[X \xrightleftarrows{i}[p] X \oplus Y \xleftrightarrows{j}[q] Y\] exists, so $\cat C$ has finite coproducts;
    \item if $X_1, \ldots, X_n \in \cat C$ then the map $X_1 \cotimes \cdots \cotimes X_n \to X_1 \times \cdots \times X_n$ such that the composite
                  \[ X_i \to X_1 \cotimes \cdots \cotimes X_n \to X_1 \times \cdots \times X_n \to X_j \]
                is $\id_X$ if $i = j$ and $0$ if $i \neq j$ is an isomorphism. We write $X_1 \oplus \cdots \oplus X_n$ for the n-ary biproduct.
    \end{itemize}

\end{rmk}
% \begin{exc}
%   \todo{something about turning map into functor, additional exercises}.
% \end{exc}

\begin{defn}\label{defn:semi-additive}
A \emph{semi-additive} category is a category $\cat C$ with finite products and coproducts such that the following conditions are satisfied:
\begin{enumerate}
\item\label{defn:semi-additive:zero-object} the natural map $\initial \to \terminal$ is an isomorphism (so $\cat C$ has a zero object $0$),
\item\label{defn:semi-additive:biproduct} for $X_1, ..., X_n$ in $\cat C$ the map $X_1 \cotimes \cdots \cotimes X_n \to X_1 \times \cdots \times X_n$ is an isomorphism.
\end{enumerate}
(Note that we use \cref{defn:semi-additive:zero-object} to define the map in \cref{defn:semi-additive:biproduct}.)
\end{defn}

\begin{rmk}
While being (pre-)additive is \emph{structure} on a category, being semi-additive is a \emph{property} of a category.
\end{rmk}

\begin{lem}
  If $\cat C$ is semi-additive, then it is canonically enriched in commutative monoids. If $\cat C$ was additive, then this agrees with the given enrichment in abelian groups (under the inclusion $\catAbelianGroup \hookrightarrow \catCommutativeMonoid$).
\end{lem}
\begin{proof}[name={sketch, in which the commutativity of the many diagrams is left to the reader}]
  For $f, g\colon X \to Y$ in $\cat C$, define $f + g$ to be
  \[
    X \xrightarrow{\Delta} X \times X = X \oplus X \xrightarrow{f \oplus g} Y \oplus Y = Y \cotimes Y \xrightarrow{\nabla} Y
  \] where $\Delta$ is the diagonal map and $\nabla$ is the codiagonal.

  \emph{Associativity}: For $f, g, h\colon X\to Y$, commutativity of 
\[\begin{tikzcd}
  & {X\oplus X} && {Y\oplus Y} \\
  X & {X\oplus X\oplus X} && {Y\oplus Y\oplus Y} & Y \\
  & {X\oplus X} && {Y\oplus Y}
  \arrow["\Delta", from=2-1, to=1-2]
  \arrow["{\id_X \oplus \Delta}"', from=3-2, to=2-2]
  \arrow["{\Delta \oplus \id_X}", from=1-2, to=2-2]
  \arrow["\Delta", from=2-1, to=2-2]
  \arrow["\Delta"', from=2-1, to=3-2]
  \arrow["{f \oplus(g+h)}"', from=3-2, to=3-4]
  \arrow["{(f+g)\oplus h}", from=1-2, to=1-4]
  \arrow["{\nabla \oplus \id}", from=2-4, to=1-4]
  \arrow["{\id_Y \oplus \nabla}"', from=2-4, to=3-4]
  \arrow["\nabla"', from=3-4, to=2-5]
  \arrow["\nabla", from=1-4, to=2-5]
  \arrow["{f\oplus g\oplus h}"{description}, from=2-2, to=2-4]
\end{tikzcd}\]  
  shows that $(f+g) + h = f + (g + h)$.

  \emph{Unitality}: For $f\colon X \to Y$, commutativity of 
\[\begin{tikzcd}
  & {X\oplus X} && {Y\oplus Y} \\
  X & {X \oplus 0} && {Y\oplus 0} & Y
  \arrow[from=2-1, to=2-2]
  \arrow["\id\oplus0", from=1-2, to=2-2]
  \arrow["\Delta", from=2-1, to=1-2]
  \arrow["{f\oplus \id}"', from=2-2, to=2-4]
  \arrow["{f\oplus 0}", from=1-2, to=1-4]
  \arrow["\nabla", from=1-4, to=2-5]
  \arrow[from=2-4, to=2-5]
  \arrow["{\id \oplus 0}", from=2-4, to=1-4]
\end{tikzcd}\]  
   shows $f + 0 = f$ and likewise $ 0 + f = f$.

   \emph{Commutativity}: For $f,g\colon X \to Y$, commutativity of 
\[\begin{tikzcd}
  & {X\oplus X} && {Y\oplus Y} \\
  X &&&& Y \\
  & {X\oplus X} && {Y\oplus Y}
  \arrow["\nabla", from=1-4, to=2-5]
  \arrow["\nabla"', from=3-4, to=2-5]
  \arrow["{g \oplus f}"', from=3-2, to=3-4]
  \arrow["{f\oplus g}", from=1-2, to=1-4]
  \arrow["\Delta", from=2-1, to=1-2]
  \arrow["\Delta"', from=2-1, to=3-2]
  \arrow["{\text{swap}}"{description}, from=1-2, to=3-2]
  \arrow["{\text{swap}}"{description}, from=1-4, to=3-4]
\end{tikzcd}\]   gives $f + g = g + f$.

So far, we have proved that $\Hom_{\cat C}(X,Y)$ is a commutative monoid.

\emph{Distributivity}: For $f,f'\colon X \to Y$ and $g,g'\colon Y  \to Z$, commutativity of 
\[\begin{tikzcd}
  X & {X\oplus X} & {Y\oplus Y} \ar[d, "g\oplus g"] & Y \\
  && {Z\oplus Z} & Z
  \arrow["{gf\oplus gf'}"', from=1-2, to=2-3]
  \arrow[from=1-3, to=2-3]
  \arrow["\nabla", from=1-3, to=1-4]
  \arrow["g", from=1-4, to=2-4]
  \arrow["\nabla"', from=2-3, to=2-4]
  \arrow["{f\oplus f'}", from=1-2, to=1-3]
  \arrow["\Delta", from=1-1, to=1-2]
\end{tikzcd}\]
shows that $g(f+ f') = gf + gf'$ and likewise $(g + g')f = gf + g'f$.

This finishes the proof that \(\cat C\) is enriched in commutative monoids.

\emph{Agreement with abelian enrichment}: If $\cat C$ is additive and $f,g\colon X \to Y$, then let
  \[
      X \xrightleftarrows{i_1}[p_1] X \oplus X \xleftrightarrows{i_2}[p_2] X
    \] be the biproduct, and likewise for $Y$ (by abuse of notation, we use the same names for the injection and projection maps). Then commutativity of
\[\begin{tikzcd}
  X & {X\oplus X} & X & X & {X\oplus X} & X \\
  Y & {Y\oplus Y} & Y & Y & {Y\oplus Y} & Y
  \arrow["f"', from=1-1, to=2-1]
  \arrow["{p_1}"', to=1-1, from=1-2]
  \arrow["{p_2}", from=1-2, to=1-3]
  \arrow["g", from=1-3, to=2-3]
  \arrow["{p_2}"', from=2-2, to=2-3]
  \arrow["p_1", from=2-2, to=2-1]
  \arrow["{f\oplus g}"{description}, from=1-2, to=2-2]
  \arrow["f"', from=1-4, to=2-4]
  \arrow["{i_1}"', from=2-4, to=2-5]
  \arrow["{i_1}", from=1-4, to=1-5]
  \arrow["{i_2}"', from=1-6, to=1-5]
  \arrow["{i_2}", from=2-6, to=2-5]
  \arrow["g", from=1-6, to=2-6]
  \arrow["{f\oplus g}" description, from=1-5, to=2-5]
\end{tikzcd}\]
gives $f \oplus g = (f \oplus g) \circ \id_{X \oplus X} = (f \oplus g) \circ (i_1 p_1 + i_2 p_2) = i_1 f p_1 + i_2 g p_2$, and likewise $f \oplus 0 = i_1 f p_1$, $0 \oplus g = i_2 g p_2$, so \[
      f \oplus g = (f \oplus 0) + (0 \oplus g).
    \] Then bilinearity of composition shows that the composite
  \[
      X \xrightarrow{\Delta} X \oplus X \xrightarrow{f \oplus g} Y \oplus Y \xrightarrow{\nabla} Y
    \] is $f + g$ (in the enriched sense).
\end{proof}
% What we have learnt from this long proof is that being an additive category is a \emph{property}.
\begin{cor}
  Being additive (i.e. admitting an additive structure) is a property.
\end{cor}
\begin{proof}
  The category $\cat C$ is additive if and only if if it is semi-additive and all hom-monoids are groups.
\end{proof}

\begin{exmp}
The category $\catCommutativeMonoid$ of commutative monoids is semi-additive (omitted), but not additive since $\Hom(\bb N, \bb N) = \bb N$ does not have inverses.
\end{exmp}

\section{Abelian categories}
\begin{exc}
  If $\cat C$ is pre-additive, show that \[
      \eq\bigl(\!\!\begin{tikzcd} X \ar[r, shift left, "f"] \ar[r, shift right, "g"'] & Y \end{tikzcd}\!\!\bigr) = \eq\bigl(\!\!\begin{tikzcd} X \ar[r, shift left, "f-g"] \ar[r, shift right, "0"'] & Y \end{tikzcd}\!\!\bigr) \eqcolon \ker(f-g)
    \] (if one of them exists).
\end{exc}
So an additive category $\cat C$ has finite limits or colimits if and only if it has respectively kernels or cokernels.
\begin{defn}
  A \emph{pre-abelian} category is an additive category with kernels and cokernels.
\end{defn}
\begin{defn}
  Let $\cat C$ be a pre-abelian category and $f\colon X \to Y$ in $\cat C$. The \emph{image} of $f$ is $\im(f) \coloneq \ker(Y \to \coker(f))$. The \emph{coimage} of $f$ is $\coim(f) \coloneq \coker(\ker(f) \to X)$.
\end{defn}
Warning: the zoo of names will get worse.
\begin{lem}
  Any $f\colon X \to Y$ in a pre-abelian category factors uniquely via \[
      X \to \coim(f) \to \im(f) \to Y\text{.}
    \]
\end{lem}
\begin{proof}
  The composition $\ker(f) \to X \to Y$ is $0$. So there is a unique factorisation \[
      X \to \coim(f) \to Y
    \] of $f$. The composition $\coim(f) \to Y \to \coker(f)$ is $0$ (since this holds after precomposition with $X \to \coim(f)$). So there is a unique factorisation \[
      \coim(f) \to \im(f) \to Y\text{.} \qedhere
    \]
\end{proof}
And now the moment we have all been waiting for: the definition of an abelian category.

\begin{defn}\label{defn:abelian-category}
  An \emph{abelian} category is a pre-abelian category $\cat A$ such that the induced map $\coim(f) \to \im(f)$ is an isomorphism for all $f\colon X \to Y$ in $\cat A$.
\end{defn}
\begin{exmp}
The category $\catAbelianGroup$ of abelian groups and more generally the category $\catModule_R$ of (left or right) modules over a ring \(R\) are abelian categories, and the coimage--image isomorphims in these categories is the first isomorphism theorem: $X/\ker(f) \cong \im(f)$.
\end{exmp}

Correspondingly, we also call the condition of \cref{defn:abelian-category} `the first isomorphism theorem' in general pre-abelian categories.

The following lemma tells us that the category of presheaves of abelian groups on a topological space is an abelian category. 
\begin{lem}
If $\cat A$ is abelian and $\cat C$ is small, then $\catFunctor(\cat C\opp, \cat A)$ is abelian.
In particular, the category \(\catAbelianPresheaf(X)\) of abelian presheaves on a space \(X\) is abelian.
\end{lem}
\begin{proof}
  If $\alpha, \beta\colon F \implies G$ is a natural transformation, then define the sum $\alpha + \beta\colon F \implies G$ by $(\alpha + \beta)_U = \alpha_U + \beta_U$. This is natural since the maps $F(V) \to F(U)$ and $G(V) \to G(U)$ are group homorphisms for any map $U \to V$ in $\cat C$. This turns $\catFunctor(\cat C\opp, \cat A)$ into a pre-additive category. All the other questions (existence of finite, limits and colimits, the first isomorphism theorem) are checked objectwise.
\end{proof}
\end{document}
